### 2016-10-13 paper.tex
some updates

\documentclass[nocopyrightspace]{sigplanconf}

\usepackage{amsmath}
\usepackage{color}
\usepackage[svgnames,x11names]{xcolor}
\usepackage[hyphens]{url}
\usepackage[citecolor=DarkGreen,linkcolor=DarkGreen,urlcolor=DarkGreen,colorlinks]{hyperref}
\usepackage{epsfig}

\newcommand{\TODO}[1]{\textbf{[TODO: #1]}}

\begin{document}
\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\title{Conex - establishing trust into data repositories}

\authorinfo{Hannes Mehnert}{University of Cambridge}{}
\authorinfo{Louis Gesbert}{OCamlPro}{}

\date{}

\maketitle

\begin{abstract}
  Opam is a software update system, responsible for discovering, downloading, building, and installing packages from community repositories.
  The data and metadata at each step should be authenticated to originate from a (transitively) trusted author.
  Opam does not include any mechanism to authenticate any step at the moment.
  We propose \emph{conex}, which establishes digital signatures from the package author to the user.
  Using conex, neither the distribution server nor the transport layer need to be trusted.
  Clients will install authenticated packages, and notice conex only if signatures cannot be verified.
  Authors need to adapt their workflow slightly, an updated \emph{opam publish} will be developed.
  Janitors (repository maintainers) need to adapt their workflow to emit cryptographic signatures to the repository (in addition to pressing merge PR buttons).
\end{abstract}

\section{Introduction}
Opam discovers, downloads, builds, and installs OCaml packages from community repositories.
It is a software update system which relies on a central server which distributes a repository (package metadata: dependencies and build instructions) of packages.
When a new package is released, it is inserted into the repository and distributed via the central server to all clients pulling it.
The central server is an online service and a single point of failure.
Recently those servers have been under attack \cite{adobe,apache2,apache1,opera,kernel,cabal,kernelorg,debian,freebsd,php3,savannah,maven,savannah2,npm,php,phpmyadmin,php2,github,ocamlforge,redhatssh,gentoo,rubygems,fedora,wordpress} to inject malware into packages.

So far there has been no breach of opam detected, but other package managers have been compromised, see \cite{cabal,maven,npm,rubygems}.
Opam does not provide any security mechanism at any step:
Discovery in current opam (1.2.2) is done via a https download (for the main opam repository), but the download tool (\texttt{wget}) is used with the \texttt{--no-check-certificates} command line option, disabling any certificate validation.
Thus, a person-in-the-middle can inject arbitrary metadata (additional patches, modified packages, modified checksums, ...) during the discovery step.
A package is downloaded using the URL specified in the metadata of the package (usually a secure channel using https, again without certificate validation), and checks the recorded MD5 digest (which is known to be weak~\cite{rfc6151}) against the downloaded archive.
The build step is not contained into a temporary chroot environment, but may modify arbitrary files on disk.
The recipe for building is part of the metadata, and may also include invoking any shell commands (such as downloading more files and executing those).
The install step may also modify arbitrary files.

Opam 2.0 is not released yet, but fixes major concerns:
downloading validates the TLS certificates~\cite{opampr} using system-wide trust anchors,
a wrap command is available for building and installing packages~\cite{wrappr}, optionally enabling sandboxing.
Nevertheless, a breach of the central repository server enables an attacker to modify packages arbitrarily.
Relying only on TLS for packages is suboptimal, because many TLS implementations suffer from a long history of security vulnerabilities (e.g. goto fail, Heartbleed).

In this paper we present \emph{conex}, a system which establishes digital signatures - the original author signs their package release and build instructions, the client verifies this signature - to re-establish trust into packages installed by opam.
Neither the the server hosting the repository nor the server where the tarballs are downloaded from need to be trusted!
Conex has been designed with opam in mind, but it can be used without opam to publish and update any data in an authenticated way.
Conex is not limited to the main opam repository, it can be used with any public or private opam repository.
The only requirement is a second channel to distribute an initial set of trust anchors (e.g. an opam package in the main opam repository).
The main goal of conex is to \emph{establish an authenticated way to distribute packages}.

The main opam repository is currently a git repository hosted on GitHub.
Conex does not depend on GitHub, or even on git.
Conex is constrained by the current use of the main opam repository:  most package authors release a package and submit a pull request (PR).
Janitors (repository maintainers) then check the PR, TravisCI builds it using several OCaml and opam versions, and the automated bot Camelus~\cite{camelus} reports opam inconsistencies.

Conex is implemented in OCaml and licensed under the 2-clause BSD license (source~\cite{conex}).
To provide authenticity for all other packages, conex is best installed together with opam.
This leads to a constraint on conex: it may not have any dependencies apart from opam and the OCaml runtime.
The installation of OCaml, opam, and conex need to be authenticated via other means (such as the OS package manager).

Earlier work on securing software update systems, such as the update framework (TUF)~\cite{tuf} focusses on distributing directories securely, but due to centralised files simultaneous updates of packages are not free of conflicts.
Diplomat~\cite{diplomat} was designed at a similar time as conex, and shares various goals as conex.
%XXX: there need to be a more concrete comparison with diplomat somewhere
Our earlier proposal for opam signing~\cite{opamsigning} introduced unnecessary structure and complexity by adjusting TUF to the decentralised opam workflow.

In section~\ref{sec:design} we discuss the design of conex, followed by a timestamping service in section~\ref{sec:shortcuts}.
Afterwards in section~\ref{sec:implementation} we briefly present the implementation of conex and opam integration.
Related work is mentioned in section~\ref{sec:related}, we conclude in section~\ref{sec:conclusion}.

\section{Design} \label{sec:design}
Conex does not claim to protect against undiscovered vulnerabilities in the packages, dependent tools, or conex itself.
It is also not a scanner for installed vulnerable packages.

\subsection{Threat model}
There are many risks that users of software update systems face, ranging from injecting traffic over weaknesses in TLS, weaknesses in the network infrastructure, compromising signing keys, to compromising the repository or signing infrastructure.

This leads us to consider a threat model where some involved systems are compromised.  We assume that an attacker can:
\begin{itemize}
  \item Compromise the central server distributing the repository.
  \item Respond to user requests (acting as person-in-the-middle).
\end{itemize}

An attack is successful if the attacker can change the contents of a package that a user installs, or preventing a user from updating to the most recent version of a package.

% LG: What about injecting packages (dependencies) into the user system ? I have
% \texttt{provides:} in mind, that may enable this without changing the contents
% of a package

%(offline!) keys -- we need to be precise which keys are where..
Conex cannot detect compromises of keys.  But if janitors, authors, and other people carefully read through updates of packages, releases not done by the original author will be detected.
We assume that authors and janitors protect their private keys on their computers in a reasonable way, but we also have mechanisms in place for key revocation.
If a private key is compromised, we consider conex security to be effective if the impact is limited to the authorised packages of the compromised key.

\subsection{Roles}
There are two roles in conex:
\begin{itemize}
  \item An \emph{author} who develops packages and releases them to the repository.  Each author signs their own packages.
  \item A \emph{janitor} who keeps the repository in a working state (fix up reverse dependency constraints etc.).  Janitors also maintain the package name and author name resources by signing them.  Instead of trusting a single janitor to sign these resources, a quorum of signatures of janitors is required to avoid this concentration of power.
\end{itemize}

A janitor may also act as an author, in case they develop and maintain packages on their own.  There is no need for a janitor to have multiple private keys.
The set of janitors is a subset of the authors.

\subsection{Resources}
We distinguish various resources which are all hosted in the same repository:
author information (public key, id, contact information), teams (members), package metadata (build recipe, dependencies, url to tarball), list of releases (of a single package), and authorisation (who is maintaining each package).
For each author one signed index file is in the repository, containing all resources and digests the author vouches for.
A janitor is a member of the team ``janitors''.

Each resource is only valid if it is vouched for either by an authorised key (in case of relesaes and package checksums), or a quorum of janitors.
Each resource contains a counter, monotonically increasing between updates, to prevent rollback attacks.
A quorum of janitors can reset the counter to recover from fast forward attacks.

There are two universes of names which need to be protected against starvation attacks (and typographic ones~\cite{typo}): package names and key identifiers.
Janitors need to carefully check new package and public key submissions.

Each author updates their own packages, which should be free of conflicts between concurrent repository updates, due to separate resources for each package.
% unclear without VC in mind, which is'nt introduced yet: precise ``conflicts between concurrent repository updates'' ?
If a conflict is detected, it is up to the janitors to decide which update should go through, and which should be resubmitted on top of the other.
%Modifications to team membership can only be done sequentially 
%The main source of conflict is likely team membership addition and deletion.

\subsection{Repository layout}
Conex is not bound to a file system, but we use their terminology.
The root directory has three subdirectories, \texttt{keys}, \texttt{index} and \texttt{packages}.

The \texttt{keys} directory contains one file for each author and each team, named after their identifier.
The \texttt{index} directory contains one index file for each author, named after their key identifier.
The \texttt{packages} directory consists of subdirectories, which each must contain an \texttt{authorisation} file and a \texttt{releases} file, plus possibly any subdirectories.
Each subdirectory must be listed in the \texttt{releases} file, and must contain a \texttt{checksum} file besides any other data files and further subdirectories (all listed in the checksum file).

Each file contains a set of key-value pairs.
Any key-value representation which can be non-ambiguously normalised into a string, can be used (such as the opam format, json, s-expressions).
Each index file contains a \texttt{signed} structure: \texttt{(signed: }\textbf{resource}\texttt{, sigs: }\textbf{(id, signature, timestamp) list}\texttt{)}.
The amount of public key operations for verification is at most the number of authors of the repository.
The different resources are explained in more detail below.
The signature is computed over the normalised string representation of the resource concatenated with the public key id, and the current timestamp, using a space character as separator.

\begin{figure}[h]
\textbf{id}: restricted string\\
\textbf{name}: restricted string\\
\textbf{timestamp}: Seconds since UNIX epoch as \textbf{Int64}\\
\textbf{signature}: Base64 encoded RSASSA-PSS with SHA256 (PKCS1)\\
\textbf{public key}: RSA public key ($\geq$ 2048 bit, PKCS1, PEM encoded)\\
\textbf{service}: \texttt{Email} of \textbf{name} $|$ \texttt{GitHub} of \textbf{name} $|$ \texttt{Other} of \textbf{id} * \textbf{name}\\
\textbf{digest}: SHA256 digest of file\\
\textbf{byte size}: size as \textbf{Int64} of file\\
\textbf{resource type}: PublicKey $|$ Team $|$ Checksum $|$ Releases $|$ Authorisation
\caption{\label{fig:datatypes0}
  Data types for conex version 0}
\end{figure}

The data types for version 0 of conex are presented in Figure~\ref{fig:datatypes0}.
A \texttt{restricted string} is a case insensitive 7 bit ASCII string without control or white space characters.
Depending on usage, more characters are forbidden (e.g. opam package names may not contain a '.').
Distinction of id and name is only for clarity (ids are public key identifiers, names are package names), they form two disjoint sets of identifiers.
Upon insertion of new identifiers, checks for non-collision and validity are done.

Each file includes a version (0 for this proposal) and a monotonic 64 bit counter (starting at 0).

\begin{figure}[h]
  \begin{minipage}{\hsize}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
key: \textbf{public key},\\
identifier: \textbf{id},\\
accounts: \textbf{service list}
 \caption{\label{fig:keyformat} Public key file}
    \end{minipage}
    \hspace*{.5cm}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{id},\\
members: \textbf{id list}
 \caption{\label{fig:teamformat} Team file}
    \end{minipage}
  \end{minipage}
\end{figure}

The file format for a public key, presented in Figure~\ref{fig:keyformat}, contains the PEM encoded public key, the identifier (used for signatures and authorisations in opam), and accounts in other systems.

A level of indirection is a team, shown in Figure~\ref{fig:teamformat}, which contains a team name and its members.  A package can be owned by public keys and teams.

\begin{figure}[h]
  \begin{minipage}{\hsize}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{name},\\
authorised: \textbf{id list}
 \caption{\label{fig:authformat} Authorisation file}
    \end{minipage}
    \hspace*{.5cm}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{name},\\
releases: \textbf{name list}
 \caption{\label{fig:releasesformat} Releases file}
    \end{minipage}
  \end{minipage}
\end{figure}

The authorisations file format, shown in Figure~\ref{fig:authformat}, contains the package name and a list of authorised identifiers, either refering to public keys or teams.

Each releases file (Figure~\ref{fig:releasesformat}) contains the package name and a list of all releases.
The purpose is to avoid rollback attacks on an individual package.
It is signed by an authorised author or a quorum of janitors.

\begin{figure}[h]
  \begin{minipage}{\hsize}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
name: \textbf{name},\\
files: \textbf{(name, size, digest) list}
 \caption{\label{fig:checksumformat} Checksum file}
    \end{minipage}
    \hspace*{.5cm}
    \begin{minipage}{.4 \hsize}
counter: \textbf{Int64},\\
version: \textbf{Int64},\\
identifier: \textbf{id},\\
resources: \textbf{(name, resource type, digest) list}
 \caption{\label{fig:janindexformat} Index file}
    \end{minipage}
  \end{minipage}
\end{figure}

The checksums file format, shown in Figure~\ref{fig:checksumformat}, contains the release name, and a list of file names which are part of the release (opam, optionally patches, ...), their sizes and digests.
This list must contain all files recursively present in the current directory (excluding the checksum file).

The index file format, shown in Figure~\ref{fig:janindexformat}, contains the identifier, and all resources the author vouches for.
It may contain any checksums, even of resources not yet in the repository (but proposed in some pull request).
This avoids merge conflicts if an author wants to vouch for multiple versions of the same resources at the same time.
%\TODO{to avoid conflicts: it may contain checksums of resources not yet in the repository (in this way, multiple PRs may exist, which are merged once quorum is reached, but individual janitors can put the checksums into their index upfront!)}

\subsection{Chain of Trust}
A package is valid if and only if the chain of trust can be verified.
All files in the subdirectory are registered in the checksum file, which is signed by either an author (listed in the authorisation file in the parent directory) or a quorum of janitors.
The subdirectory name must be present in the valid releases file of the parent directory.
The authorisation file itself has to be signed by a quorum of janitors.
Each public key has to be signed by its private key and a quorum of janitors.

To bootstrap the chain, we assume an existing set of janitors (which cardinality is above the quorum).
The set for the main opam repository is distributed with opam.
The opam installation, including the initial set of janitors, is out of scope for conex (e.g. by another package manager, OpenPGP signatures).

\subsection{Verification}
Since these verification steps are done by automated systems (see section~\ref{sec:shortcuts}) as well, verification errors are not expected at the client and treated as fatal.
The user is informed about the potential security breach, the set of patches are not applied locally, and the user is asked to inform the janitors via email.
Certainly, the user can try again to update their repository.

A client can verify a repository in two ways:
\begin{itemize}
  \item a \emph{complete snapshot} of the repository (fresh clone)
  \item an already verified repository and a list of \emph{updates} (in the form of patches) which should be applied
\end{itemize}

We will briefly outline the algorithm for both subsequently.

\paragraph{Snapshot verification}  Verification of a snapshot is straightforward: take the set of initial janitor keys as trusted, verify all index files from this set, verify the janitors team, iterate over all packages, verify the authorisation file (properly signed by a quorum of janitors), verify all releases files, and verify all checksum files.
The client also needs to ensure that there are no subdirectories or files which are not listed in the releases or checksum files.
The public keys are verified upon demand: whenever a signature is encountered, the key of the signing id is verified.

\paragraph{Update verification}  The repository on disk is already verified.
The client uses it to verify each update separately.
Each modified resource must be signed, either by an authorised author or by a quorum of janitors.
The counter must start at 0 (upon creation) and increment.
The file name and identifier must match.
A key rollover needs to introduce the new key (replacing the old), and a signature with the new key of the index file.
Removal of resources is done by replacing it with the empty string.
A new package needs to have an authorisation file (signed by a quorum) and a releases file (signed by an authorised author).
%% \begin{itemize}
%%   \item janitor index modification
%%   \item janitor index creation - counter 0, properly named, signed by janitor
%%   \item janitor index deletion - counter increasing, properly named, signed by quorum
%%   \item key modification -- key is sufficiently big, unique id matching filename, counter is increased, signed by quorum, no signature with old key
%%   \item key insertion -- key is sufficiently big, unique id matching filename, counter is 0, signed by quorum
%%   \item key deletion -- counter increasing, no signatures with this key, empty string is used to denote no key, signed by quorum
%%   \item role modification -- counter increasing, no signature with old role, signed by quorum
%%   \item authorisation insertion -- counter is 0, id matches directory, signed by quorum, authorised key ids are present
%%   \item authorisation modification -- counter increasing, all checksum are signed with an authorised key, signed by quorum
%%   \item complete package removal -- empty authorised, empty releases, signed by quorum
%%   \item package insertion -- valid authorisation, releases, and checksum, both counter 0, signed by author (authorisation by quorum)
%%   \item package update -- valid checksum, counter increasing, signed by author
%%   \item new package version release -- index counter increasing and new version added to list, valid checksum, signed by author
%%   \item package version deletion -- counter increasing, removed from releases, signed by author
%% \end{itemize}

\subsection{Security Analysis}
%\TODO{from opam signing: When a developer key is stolen, the snapshot key is also needed to target a single user for a malicious patch.}
%\TODO{still sort of a laundry list}

If package metadata and public keys are distributed from different sources, an attacker who compromised a key could hinder a user from receiving the key revocation by preventing communication to the key server.
This would allow the attacker to continue signing packages with the compromised key and having the user install arbitrary data.
To avoid this, package metadata, signatures, and public keys are in the same repository.
%Package metadata, signatures, and public keys are distributed in the same repository to prevent attackers from hindering one communication link, e.g. the public key one.
%This would otherwise allow a detected attacker who compromised a key to deliver metadata signed with this compromised key.
%% LG: I am not sure I really get your meaning here
% HM: hopefully now it is better!?

Each resource includes a monotonically increasing counter to prevent rollback attacks.
Otherwise, an attacker could replay e.g. an insecure package build recipe.
To prevent fast forward attacks, a quorum of janitors can reset counters.
Also, each resource includes a name, otherwise the signature of package A would be valid for package B.
Each signature contains the resource type to avoid carrying over a signature from one resource to another.

Instead of removing resources to revoke them, the already claimed names stay there.
The public key or package is emptied and must be signed by a quorum of janitors.

A person-in-the-middle attacker could implement a backdoor for a package, but would need to either modify the checksum, prepare a new release, embed the patch as metadata in the repository, or introduce a new package and add a dependency.
All of these actions require either a signature from the original author or from a quorum of janitors.
Any suspicious behaviour in widely used packages will likely be spotted by the community which reads through the commit logs.

An attacker can freeze a client from updating (by preventing the client to communicate with the repository server), but downgrades of packages are not possible (an update cannot remove packages without appropriate signatures).

Conex does not depend on unreliable system time and comparison of timestamps, instead it uses monotonic counters.
Other systems rely on git signing, which is out-of-tree data and thus an attacker can prevent a user from receiving this data (while still receiving (potentially malicious) updates).

In the case conex cannot verify package metadata, it will not pass this unverified data to opam.
In case conex cannot verify a downloaded tarball, this will be reported.

Mix-and-match attacks provide e.g. outdated releases of package A and B, up-to-date releases of package C, are not mitigated in conex since there is no central list of releases.
They can be mitigated by relying on some append-only log, blockchain, or other monotonic updates, such as git without forced pushes.
We will discuss a timestamping service in the next section, mitigating both mix-and-match and freeze attacks.

\section{Timestamping Service} \label{sec:shortcuts}
%\TODO{see also http://bentnib.org/posts/2016-04-12-authenticated-data-structures-as-a-library.html}

%In the presented setup, conex requires verification of at least 8 signatures (3 for the public key, 3 for the authorisation, then index and checksum) per package.
%Valid public key can be cached during processing, leaving the number at 5, which is still too high, especially for embedded systems.

To mitigate mix-and-match attacks, a \emph{timestamp service}, which updates the repository regularly, verifies all updates, and attaches its digital signature to the latest patch.
The public key of the timestamp service is part of the repository and signed by multiple janitors.
If there are multiple instances of a timestamp service run by different organisations, and its code is well audited, it might be reasonable to only verify their signatures, and trust it to verify all author signatures.
While there is not yet an implementation of a timestamp service, it will likely be implemented as a MirageOS~\cite{mirage,nqsb-tls} application, thus the trusted code base is small.

The main opam repository served as an example over the last section, but conex can easily be applied to any other opam repository.
Initial set of janitors (can be a single element), the quorum (can be one) must be distributed for each other repository.
Of course, an automated timestamp service can be used for other repositories as well.

%\TODO{this is insecure, since an attacker can add themselves to the authorisation file and self-sign it!}
If the work overhead for janitors turns out too high, we can leverage the system in a way that \emph{neither authorisation nor public keys require a quorum of janitors}.
Diplomat~\cite{diplomat} introduced online keys for this (their \texttt{new} role).
This will open the system to name squatting attacks.
Snapshot verification does not apply anymore, because there is no history of authorisations.
Update verification can still be done, thus we could trust the timestamp notaries and use git.

%that they contain the same information as the original ones, but updates can be verified (and again, we can outsource trust into the timestamping service).

Another useful automated system will be a \emph{PR bot}, similar to Camelus, which verifies each PR on the main opam repository and reports the results of the update (whether it is valid, how many janitors need to sign, etc.).

Automated systems which build a package and record the (platform-dependent) digests of the resulting binaries can be integrated to sign their results in the repository.
Since OCaml 4.03 produces reproducible binaries, thus the locally produced binaries can be checked.
Conex can also serve for binary distribution.

\section{Conex implementation} \label{sec:implementation}
%\TODO{UX: timestamp in signature!}

Conex is work-in-progress~\cite{conex}.
It will be abstracted over the data layout (json, opam, s-expression), and over a cryptography provider.
One provider will be nocrypto, another the \emph{openssl} utility, enabling bootstrapping without opam.
The timestamp service and other automated tools will use the conex library.

Conex is not yet integrated into opam, but work on opam 2.0, such as compilers-as-packages, are prerequisites for conex.
The opam publish utility will support authors: creating private keys, creating checksum, authorisation, and releases files, and signing these.
There is a demand for tooling for janitors, such as a queue of updates requiring signatures for a quorum.

\section{Related work} \label{sec:related}
Conex is based on the update framework~\cite{tuf}, which centralises all packages into a single \texttt{target.txt} file.
Instead of doing this, conex was designed with distributed updates in mind.

More recent is Diplomat~\cite{diplomat}, which automatically promoting unclaimed packages to claimed (which we can do using timestamp service).
Our security model instead uses digital signatures of packages, and relies on authors to sign their packages.

The Haskell signing proposal does not enforce authors to sign their packages, but the repository is signed with an online key.

Our original opam signing proposal~\cite{opamsigning} did not include a list of releases per package, opening up to various attacks.
Additionally, our original proposal introduced an unnecessary hierarchical structure.
It relied on the timestamping service or using git as update protocol with appropriate protection of the transport layer.

%Conex can also be viewed as a verified append-only log, similar to a blockchain.  Further research is needed to

\section{Conclusion} \label{sec:conclusion}
We presented conex, which secures the software update system opam with minor modifications to the current workflow.
Conex uses signatures from package authors to users, removing the demand to trust a central server.
Conex mitigates known practical attacks as well as theoretical ones described in literature.

Future work includes finishing the implementation of conex, integrating it into opam, and developing automated services (timestamp service, build bot, PR bot).
A first beta version of conex, including opam integration, is expected to be available for the OCaml workshop 2016.

\paragraph{Acknowledgements} Parts of this work were supported by EPSRC Programme Grant EP/K008528/1 (REMS - Rigorous Engineering of Mainstream Systems).  We thank especially Joe and Rapha\"el for their feedback on earlier drafts.

\bibliographystyle{acm}
\bibliography{paper}

\end{document}